Density-dependent selection in evolutionary genetics: a lottery model of Grime's triangle

### 2016-12-21 manuscript.tex
latest

Each iteration, adults produce $m_i$ new offspring (henceforth called ``propagules''). We assume adults cannot be ousted from occupied territories, so only propagules landing on unoccupied territories are included in $m_i$. Dispersal is random and unrestricted; each propagule has an equal probability of landing on any of the $U$ unnocupied territories. Loss of propagules during dispersal is subsumed into $m_i$ ($m_i$ only counts survivors which go on to contest the territory they land in), and we assume that propagules disperse independently of each other. In general, $m_i$ will increase with $n_i$, and will also depend on population density $N$. For example, if $b_i$ is the number of propagules produced per individual from genotype $i$, then $m_i=b_i(1-N/T)n_i$ would reflect the loss of propagules that disperse to occupied territories. Nevertheless, we will assume $m_i=b_i n_i$, meaning that all propagules land on unoccupied territories (a form of directed dispersal). This choice simplifies the mathematics without seriously restricting the generality of our analysis, since the results presented here are not sensitive to the specific functional form of $m_i$.

@@

The propagule numbers $x_i$ in different territories obey a multinomial distribution with

are not independent random variables.

For sufficiently large $K$, holding $n_i/T$ fixed, the Poisson limit theorem implies that the binomial distributions for $x_i$ at each successive stage of this procedure are all closely approximated by a Poisson distribution with mean $l_i$, where we have used the fact that large $T$ implies large $U$ except in the biologically uninteresting case that there is vanishing population turnover $d_i \sim 1/T$.

>>>>

Each iteration, adults produce $m_i$ new offspring (henceforth called ``propagules''). We assume adults cannot be ousted from occupied territories, and only propagules landing on unoccupied territories are included in $m_i$. Propagules disperse at random, independently of each other, and without spatial restrictions; each propagule has an equal probability of landing on any of the $U$ unnocupied territories. Thus, there is no interaction between propagules (e.g. avoidance of territories crowded with propagules). Loss of propagules during dispersal is subsumed into $m_i$ ($m_i$ only counts propagules which go on to contest the territory they land in). In general, $m_i$ will increase with $n_i$, and will also depend on population density $N$. For example, if $b_i$ is the number of propagules produced per individual from genotype $i$, then $m_i=b_i(1-N/T)n_i$ would reflect the loss of propagules due to dispersal to occupied territories. We assume $m_i=b_i n_i$, meaning that all propagules land on unoccupied territories (a form of directed dispersal). This choice is not biologically motivated; it simplifies the mathematics without seriously restricting the generality of our analysis, since the results presented here are not sensitive to the specific functional form of $m_i$.

@@

For each genotype's dispersal, the counts of propagules across the unnocupied territories follows a multinomial distribution with equal probabilities of ``success'' for all territories. Thus, the $x_i$ in different territories are not independent random variables. However, for sufficiently large $T$, holding $n_i/T$ fixed, the Poisson limit theorem implies that this multinomial distribution for the $x_i$ accross territories is closely approximated by a product of independent Poisson distributions for each territory, each with rate parameter $l_i$. Here we have used the fact that large $T$ implies large $U$ except in the biologically uninteresting case that there is vanishing population turnover $d_i \sim 1/T$.
